A latency-tolerant runtime for large graph computations

### 2011-01-13 doc/hotpar2011/softxmt-hotpar2011.tex
comment on abstract

  Large irregular applications such as sparse, low diameter graph problems often
  perform badly on commodity clusters. These application make many
  non-sequential accesses to global data structures with low degrees of
  locality, interacting badly with commodity processors' memory
  systems, which are optimized for local, sequential memory
  accesses. Special-purpose hardware optimized for these irregular
  applications has been built; the Cray XMT is one such machine. It
  uses multithreading to tolerate the latency of low-locality memory
  accesses, and while it exhibits excellent performance and energy
  efficiency for irregular applications, it performs badly on standard
  sequential applications. We believe we can have the best of both
  worlds: we argue that modern commodity processors have many of the
  features necessary \todo{Brandon: we will have to be able to list these as features in the paper or use a different phrase; my feeling is that these are aspects of the system which are not optimized for this but are sufficient to suggest potential for enough memory concurrency} to give good performance on irregular problems,
  and that a latency-tolerant runtime designed to exploit these
  features can enable speedup for irregular applications on commodity
  processors. In this paper, we describe the key features of such a
  runtime system, and what a cluster built using our runtime might
  look like. Using a preliminary single-node implementation of our
  runtime, we obtained speedups on simple list chasing benchmarks. Our results show promise for a multi-node implementation.

>>>>

  Large irregular applications such as sparse, low diameter graph problems often
  perform badly on commodity clusters. These application make many
  non-sequential accesses to global data structures with low degrees of
  locality, interacting badly with commodity processors' memory
  systems, which are optimized for local, sequential memory
  accesses. Special-purpose hardware optimized for these irregular
  applications has been built; the Cray XMT is one such machine. It
  uses multithreading to tolerate the latency of low-locality memory
  accesses, and while it exhibits excellent performance and energy
  efficiency for irregular applications, it performs badly on standard
  sequential applications \todo{Brandon: should this be more specific or worded differently? Its true sequential perform badly since switching between many threads, but also does XMT also perform  poorly (relative to alternatives) on highly parallel but regular apps with lots of temporal and spatial locality?}. We believe we can have the best of both
  worlds: we argue that modern commodity processors have many of the
  features necessary \todo{Brandon: we will have to be able to list these as features in the paper or use a different phrase; my feeling is that these are aspects of the system which are not optimized for this but are sufficient to suggest potential for enough memory concurrency} to give good performance on irregular problems,
  and that a latency-tolerant runtime designed to exploit these
  features can enable speedup for irregular applications on commodity
  processors. In this paper, we describe the key features of such a
  runtime system, and what a cluster built using our runtime might
  look like. Using a preliminary single-node implementation of our
  runtime, we obtained speedups on simple list chasing benchmarks. Our results show promise for a multi-node implementation.
